---
title: 'Homework #4'
author: "Eric Pettengill"
output: pdf_document
---

### Using the \texttt{Default} dataset do the following.

(a) Fit a logistic regression model that uses \texttt{income} and \texttt{balance} to predict \texttt{default}

```{r message=FALSE, warning=FALSE}
library(ISLR)
library(caret)

glmfit <- train(default ~ income + balance, 
             data = Default, 
             method = "glm",
             family = "binomial")
```


(b) Estimate the test error using 5-fold cross validation.

```{r}
set.seed(1018)

fitControl <- trainControl(method = "repeatedcv",
                           number = 5)

glmfit <- train(default ~ income + balance, 
             data = Default, 
             method = "glm",
             family = "binomial",
             trControl = fitControl)

glmfit$resample[,c(3,1)]

## test error
1-mean(glmfit$resample[,1])
```


(c) Add \texttt{student} to the model and estimate the test error using 5-fold cross validation. Comment.

```{r}
set.seed(1018)

glmfit <- train(default ~ income + balance + student, 
             data = Default, 
             method = "glm",
             family = "binomial",
             trControl = fitControl)

glmfit$resample[,c(3,1)]

## test error
1-mean(glmfit$resample[,1])
```


The model in (b) performs slightly better than the model in (c)

### 5.4 #6

(a) Determine the estimated standard errors for the coefficients(\texttt{income} and \texttt{balance}) of the model fit in part (a) above.

```{r}
glmfit <- glm(default ~ income + balance, data = Default, family = "binomial")

summary(glmfit)$coefficients[-1,2]
```


(b) Write a function, \texttt{boot.fn}, to use in the \texttt{boot()} function to estimate the standard errors.

```{r}
boot.fn <- function(data, index) {
  coef(glm(default ~ income + balance, data = Default, family = "binomial", subset = index))[c(2,3)]
}
```


(c) Use the \texttt{boot()} function to estimate the standard errors of the model in (a).

```{r}
set.seed(1018)

boot.obj <- boot::boot(Default, boot.fn, R = 1000)

apply(boot.obj$t, 2, sd)
```


(d) Compare the estimates to the true values.

```{r}
data.frame(Actual = summary(glmfit)$coefficients[-1,2],
           Bootstrap = apply(boot.obj$t, 2, sd))
```

The bootstrap estimate for \texttt{income} is slightly lower than the actual value given in the \texttt{glm} output while the bootstrap estimate for \texttt{balance} in slightly higher than the actual value.